ergm models
========================================================

Do this on separate page as too easy to get clashes between igraph and network packages otherwise.


```{r init, echo=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, echo=FALSE)
library(statnet)
library(stringr)

setwd("~/Big Study")
```


# Data preparation 

## Read interpersonal network data

Each person interviewed has two csv files, one has the name FirstnameLastname.csv,
the other the name FirstnameLastnameOrgs.csv.  The first has data on relationships with other invividuals, the second relationships with other organizations.  
**DO NOT PUT OTHER csv FILES IN THE DIRECTORY**

```{r readdata, cache=TRUE}
fns <- dir(pattern='*.csv')  # Gets all csv file names
# There is a file Professions.csv that has to be excluded
ix2 <- which(str_detect(fns,"Profession"))
fns <- fns[-ix2]
ix <- str_detect(fns,"Orgs")  # TRUE when filename has Orgs in it, FALSE otherwise
org.fns <- fns[ix]  # Vector of organizational data files
fns <- fns[!ix]  # Vector of interpersonal data filenames
n <- length(fns)  # number of interperson filenames
nets <- list()   # nets will hold data
for (i in 1:n){
    nets[[i]] <- read.csv(fns[i],skip=5,header=FALSE,nrows=77)
}
```

## Data is taken from each component of the list and combined.

```{r personalnets, cache=TRUE}
el <- data.frame()  # Data will be in this dataframe

for (i in 1:n){
  tmp <- nets[[i]][,c(1,2,3,4,6,7,8)]  # don't include col 5 as this is attribute
  ego <- str_sub(fns[i],1,-5)  # Take ego name from the original filename
  sp2 <- str_locate(tmp[,1],' [A-Z]')  # Locate space in alter's name
  # Then use this location to create alter name without space
  alter <- paste(substr(tmp[,1],1,sp2[,1]-1),substring(tmp[,1],sp2[,2]),sep="")
  # Correct a couple of inconsistencies in people's names
  alter <- ifelse(alter=='RachaelWiliams','RachaelWilliams',alter)  
  alter <- ifelse(alter=='SuEdwards','SueEdwards',alter)
  # Get rid of -1 s in advice variable
  tmp[,4] <- ifelse(tmp[,4]<=0, 0, 1)
  # Create data frame by stacking data from each ego
  el <- rbind(el,data.frame(ego,alter,Knows=tmp[,2],Strength=tmp[,3],
                advice=tmp[,4],leadership=tmp[,5],influence=tmp[,6]))
}
ix <- order(el$ego)
el <- el[ix,]

```

## Add attribute data

```{r attributes}
## Basic attribute data, has whether interviewed and org name
atts <- read.table("Attributes.txt",header=TRUE,stringsAsFactors=FALSE)
atts <- atts[,-3]
orgs <- sort(unique(atts[,3]))
org.ids <- 1:length(orgs)
orgs.dta <- data.frame(OrganizationName=orgs, OrganizationID = org.ids)

atts.2 <- merge(atts,orgs.dta,by="OrganizationName")
  ix2 <- order(atts.2$Name)
atts.2 <- atts.2[ix2,]

## > names(atts)
##[1] "Name"             "Interviewed"      "OrganizationName"
 
## Data on professional background
prof <- read.csv('Profession.csv',stringsAsFactor=FALSE)
prof$Name <- paste(prof[,1],prof[,2],sep='')
prof <- prof[,-c(1,2)]
prof[,1] <- factor(prof[,1],levels=c(1:5,99),labels=c('Nurse','Doctor','Allied HP','Social service','Manager','Unknown'))
ix <- prof[,1] == 'Unknown'
prof[ix,1] <- NA
all.atts <- merge(atts.2,prof,by='Name',all.x=TRUE)
names(all.atts)
# [1] "Name"             "Interviewed"      "OrganizationName" "Profession"      
```

```{r mergeorgs}
## Test 
ix.e <- order(el$ego)
el <- el[ix.e,]

if (all.equal(atts.2[atts.2$'Interviewed'=='Yes','Name'],levels(el$ego))){

  # Merge network data and attribute data
  el.2 <- merge(el,all.atts,by.x="ego",by.y='Name',all.x=TRUE)
  names(el.2)[8] <- 'EgoOrganization'
  names(el.2)[10] <- 'EgoOrganizationID'

  el.3 <- merge(el.2,all.atts,by.x="alter",by.y='Name',all=TRUE)
  names(el.3)[12] <- 'AlterOrganization'
  names(el.3)[14] <- 'AlterOrganizationID'
  ix.e2 <- order(el.3$ego)
  el.3 <- el.3[ix.e2,]
  names(el.3)
#  [1] "alter"               "ego"                 "Knows"              
#  [4] "Strength"            "advice"              "leadership"         
#  [7] "influence"           "EgoOrganization"     "Interviewed.x"      
# [10] "EgoOrganizationID"   "Profession.x"        "AlterOrganization"  
# [13] "Interviewed.y"       "AlterOrganizationID" "Profession.y"   
  
  el.org <- el.3[,c(8,12,3,5,6,7)]
  attrib.org <- el.3[,c(10,2,1,4,8,12,9)]
  names(attrib.org)[7] <- 'Interviewed'
}
```
## Create `network` objects

```{r netobj}
adv.adj <- igraph::get.adjacency(advice, sparse=FALSE)
adv.int <- igraph::V(advice)$Interviewed
adv.net <- network(adv.adj)
set.vertex.attribute(adv.net,'Interviewed',adv.int)
adv.prof <- igraph::V(advice)$Profession
set.vertex.attribute(adv.net,'Profession',adv.prof)

advice.int.net <- adv.net
delete.vertices(advice.int.net,which(advice.int.net %v% 'Interviewed' == 'No'))

lead.adj <- igraph::get.adjacency(leader, sparse=FALSE)
lead.int <- igraph::V(leader)$Interviewed
lead.net <- network(lead.adj)
set.vertex.attribute(lead.net,'Interviewed',lead.int)
lead.prof <- igraph::V(leader)$Profession
set.vertex.attribute(lead.net,'Profession',lead.prof)

lead.int.net <- lead.net
delete.vertices(lead.int.net,which(lead.int.net %v% 'Interviewed' == 'No'))
indeg.adv <- sna::degree(advice.int.net, cmode='indegree')
outdeg.adv <- sna::degree(advice.int.net, cmode='outdegree')
set.vertex.attribute(advice.int.net,'InDegree',indeg.adv)
set.vertex.attribute(advice.int.net,'Sqrt.InDegree',sqrt(indeg.adv))
set.vertex.attribute(advice.int.net,'OutDegree',outdeg.adv)
set.vertex.attribute(advice.int.net,'Sqrt.OutDegree',sqrt(outdeg.adv))

gden(advice.int.net)
gtrans(advice.int.net)
centralization(advice.int.net,betweenness)
centralization(advice.int.net,closeness)

gden(lead.int.net)
gtrans(lead.int.net)
centralization(lead.int.net,betweenness)
centralization(lead.int.net,closeness)

```

```{r ergm1}
e0 <- ergm(adv.net ~ edges + nodematch('Profession', diff=TRUE, keep=c(1,2,5)))
summary(e0)


e1 <- ergm(advice.int.net ~ edges + mutual + nodecov('Sqrt.InDegree') + nodecov('Sqrt.OutDegree') + gwesp(0.25, TRUE), control=control.ergm(MCMLE.maxit=20))
summary(e1)

e1.gof <- gof(e1,verbose=TRUE)

x11()
par(mfrow=c(2,2))
par(oma=c(0.5,2,1,0.5))
plot(e1.gof, plotlogodds=TRUE)

summary(e1.gof)

e2 <- ergm(lead.int.net ~ mutual + ctriple + ttriple + nodematch('Profession'))
summary(e2)

```
