Network Study
============================================
Basic network properties based on the first ten respondents (shown as the first ten in the tables, identified in red on the plots).

Ideas for analysis
-----------------------------------------------

- looking for sub-groups
- looking for gaps
- looking for stars and isolates
- looking at strength of ties, how does that compare
- looking at pattern of advice; can we identify knowledge brokers?
- what's the leadership structure?
- influential outside network? Different from leaders?
- are there bottlenecks?
- bridges or brokers? what and who?
- density? centrality?
- two-mode network?

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, comment='', dev='png',cache=TRUE,dpi=100,fig.width=7)
```

```{r make-network}
library(igraph)
library(stringr)
library(xtable)
library(ascii)
library(ggplot2)

fns <- dir(pattern='*.csv')
n <- length(fns)
nets <- list()
for (i in 1:n){
    nets[[i]] <- read.csv(fns[i],skip=5,header=FALSE,nrows=77)
}

     
el <- data.frame()
#i <- 1
for (i in 1:n){
  tmp <- nets[[i]][,c(1,2,3,6,7,8)]
  ego <- str_sub(fns[i],1,-5)
#   sp <- str_locate(ego,'[a-z][A-Z]')
#   ego <- paste(substr(ego,1,sp[1]) ,  substring(ego,sp[2]))
  sp2 <- str_locate(tmp[,1],' [A-Z]')
  alter <- paste(substr(tmp[,1],1,sp2[,1]-1),substring(tmp[,1],sp2[,2]),sep="")
#  ix <- tmp[,2] > 0
#  tmp <- tmp[ix,]


alter <- ifelse(alter=='RachaelWiliams','RachaelWilliams',alter)  
alter <- ifelse(alter=='SuEdwards','SueEdwards',alter)
  
# Get rid of -1 s in advice variable
  tmp[,4] <- ifelse(tmp[,4]<=0, 0, 1)
  el <- rbind(el,data.frame(ego,alter,Knows=tmp[,2],Strength=tmp[,3],
                advice=tmp[,4],leadership=tmp[,5],influence=tmp[,6]))
}


# Reduce to non-zero edges and build a graph object
el.nonzero.edges <- subset(el, (Knows > 0 | advice > 0))


g1 <- graph.data.frame(el)

alters <- V(g1)

attribs <- read.table('Attributes.txt',header=TRUE,stringsAsFactors=FALSE)
#attribs[,1] <- attribs[,1]-1
#names(attribs)[1] <- "Vertex"
# Set vertex attributes
for (i in V(g1)) {
  for (j in names(attribs)) {
     ix <- match(V(g1)$name[i],attribs[,1])
    if (is.na(ix)) stop("Invalid name")
    g1 <- set.vertex.attribute(g1, j, index=i, attribs[ix,j])
  }
}

int_vertex_colors <- get.vertex.attribute(g1,"Interviewed")
colors <- c('Lightblue', 'Red')
int_vertex_colors[int_vertex_colors == 'No'] <- colors[1]
int_vertex_colors[int_vertex_colors == 'Yes'] <- colors[2]
V(g1)$colour <- 'Lightblue'
V(g1)[V(g1)$Interviewed=='Yes']$colour <- 'Red'

```
# Summary of the complete network
```{r complete}
summary(g1)
```

# Problems with data.

1. In some data sheets Rachael Williams spelt Rachael Wiliams (only one l)
2. Sue Edwards sometimes spelt Su Edwards

# "Is this person known to you?"
## Summary of the network
```{r summary-knows}
knows <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "Knows")==0])
summary(knows)
```

## Graph of the network

```{r knows-plot, message=FALSE, fig.cap='Is person known to you', out.width='10in'}
deg.know <- degree(knows,mode='in')
plot(knows, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=deg.know/2,vertex.label=NA,vertex.color=int_vertex_colors)
```

## In-degrees

```{r know-degree,comment='',results='asis'}
deg.know <- degree(knows,mode='in')
deg.dist.know <- degree.distribution(knows,mode='in')
knows.noiso <- delete.vertices(knows,V(knows)[degree(knows)==0])
print(xtable(data.frame(Name=alters$name,InDegree=deg.know),digits=0),type='html',html.table.attributes='border=0')
deg.know.df <- data.frame(table(deg.know))
print(deg.know.df,type='html',html.table.attributes='border=0')
ggplot(deg.know.df,aes(x=deg.know,y=Freq)) + geom_bar(stat='Identity') + xlab('In degree') + ylab("Frequency")
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.know), 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

# "Would you go to this person for advice?"
## Summary of the network

```{r advice}
advice <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "advice")==0])
summary(advice)
```

```{r advice-plot, fig.cap='Would you go to this person for advice?'}
advice.noiso <- delete.vertices(advice,V(advice)[degree(advice)==0])
deg.adv.no <- degree(advice.noiso, mode='in')
plot(advice.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3, vertex.label=NA,vertex.color=V(advice.noiso)$colour,vertex.size=deg.adv.no/2)
```
## In-degrees
```{r advice-degree, comment='', results='asis'}

deg.adv <- degree(advice,mode='in')
deg.dist.adv <- table(deg.adv)
deg.dist.adv.df <- data.frame(Degree=as.numeric(names(deg.dist.adv)),Frequency=deg.dist.adv)
print(xtable(data.frame(InDegree=deg.adv),digits=0),type='html',html.table.attributes='border=0')
print(xtable(data.frame(Degree=deg.dist.adv)))

ggplot(data.frame(deg.dist.adv),aes(x=deg.adv,y=Freq))+ geom_bar(stat="Identity")+xlab("In degree")+ylab('Frequency')
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.adv), 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

# "Does person occupy leadership role?"

## Summary of the network

```{r leader}
leader <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "leadership")!=1])
summary(leader)
```

## Graph of the network

```{r leader-plot, message=FALSE, fig.cap='Does the person occupy a leadership role?'}
deg.lead <- degree(leader,mode='in')
 
leader.noiso <- delete.vertices(leader,V(leader)[degree(leader)==0])
deg.lead.no <- degree(leader.noiso, mode='in')
plot(leader.noiso,layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.label=NA,vertex.size=deg.lead.no/2,vertex.color=V(leader.noiso)$colour)

```

## In-degrees
```{r leader-degree, results='asis'}
lead.op <- data.frame(InDegree=deg.lead)
print(xtable(lead.op,digits=0),type='html',html.table.attributes='border=0')
deg.lead <- degree(leader,mode='in')
deg.lead.df <- data.frame(table(deg.lead))
ggplot(deg.lead.df,aes(x=deg.lead,y=Freq))+geom_bar(stat='Identity')+xlab('In degree')+ylab('Frequency')
#data.frame(Name=alters$name,InDegree=deg.lead)
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.lead), 1, function(X) paste(X, collapse=" | ")), sep = "\n")

```

# "Do you see the person as influential outside network?"
## Summary

```{r influence}
influence <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "influence")!=1])
summary(influence)
```

## Graph of the network (isolates removed)

```{r influence-plot, fig.cap='Is the person influential outside network?'}
deg.inf <- degree(influence, mode='in')
infl.noiso <- delete.vertices(influence,V(influence)[degree(influence)==0])
deg.inf.no <- degree(infl.noiso, mode='in')
plot(infl.noiso,layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.label=NA,vertex.size=deg.inf.no/2,vertex.color=V(infl.noiso)$colour)
```
## In-degree
```{r influence-degree, results='asis'}
print(xtable(data.frame(InDegree=deg.inf),digits=0),type='html',html.table.attributes='border=0')
deg.inf.df <- data.frame(table(deg.inf))
ggplot(deg.inf.df,aes(x=deg.inf,y=Freq))+geom_bar(stat='Identity') + xlab('In degree') + ylab('Frequency')
#data.frame(Name=alters$name,InDegree=deg.inf)
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.inf), 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

# Summary of all in-degrees
```{r summary, results='asis'}
op <- data.frame(Knows=deg.know,Advice=deg.adv,Leadership=deg.lead,Influence=deg.inf)
print(xtable(op,digits=0),type='html',html.table.attributes='border=0')
#cat("Name | Knows | Advice | Leadership | Influence ", "--- | --- |--- | --- | --- ", sep="\n")
#cat(apply(op, 1, function(X) paste(X, collapse=" | ")), sep = "\n")
#cat("\n")
```

# Community structure.
*Communities* are highly connected between them but with few links to other vertices.  Calculate walktrap community using the *advice* nextwork, with isolates removed. This uses a random walk algorithm to identify communities. Pascal Pons, Matthieu Latapy: Computing communities in large networks using random walks, http://arxiv.org/abs/physics/0512106.

```{r community, eval=FALSE}
wtrp.adv <- walktrap.community(advice.noiso,steps=5)
#wtrp.adv
memb.adv <- community.to.membership(advice.noiso,wtrp.adv$merges,steps=12)
#memb.adv
ix <- order(memb.adv$membership)
com.df <- data.frame(Name=get.vertex.attribute(advice.noiso,"name"),Community=memb.adv$membership)[ix,]
com.df
```
<!---
Modularity of this community is `r igraph::modularity(advice.noiso, memb.adv$membership)`.
-->

## Edge betweenness centrality
```{r edge-betweenness, eval=FALSE}
eb.adv <- edge.betweenness.community(advice.noiso,directed=FALSE)
eb.adv.mem <- community.to.membership(advice.noiso, eb.adv$merges, steps=9)
#modularity(advice.noiso, eb.adv.mem$membership)
#table(eb.adv.mem$membership)
#eb.adv$bridges
#eb.adv$removed.edges    
#eb.adv$edge.betweenness
n <- length(eb.adv$removed.edges)
brdg.colors <- rep('grey',n)
brdg.colors[eb.adv$removed.edges[1:5]+1] <- 'red'
brdg.wd <- rep(1,n)
brdg.wd[eb.adv$removed.edges[1:5]+1] <- 5
nv <- vcount(advice.noiso)
V(advice.noiso)$size <- 5
V(advice.noiso)[V(advice.noiso)$name=='NicolaFitzmaurice']$size <- 10

plot(advice.noiso, layout=layout.kamada.kawai,edge.arrow.size=.8, vertex.label=NA,edge.color=brdg.colors,edge.width=brdg.wd,vertex.size=V(advice.noiso)$size)
```

<!---
Plot shows the five edges that provide the most connectivity.  These are as follows:
-->
```{r important-edges, eval=FALSE}
E(advice.noiso)[eb.adv$removed.edges[1:5]]
```
# Network characteristics
Have to restrict to ties between respondents only.

```{r densityfunction,echo=TRUE}
 dens <- function(g){
   n.edge <- ecount(g)
   n.vert <- vcount(g)
   e.max <- n.vert^2 - n.vert
   if (is.directed(g)){
     n.edge/e.max
   }
   else {
     n.edge/(e.max/2)
   }
 }
```

```{r interview-networks}
knows.int <- delete.vertices(knows,V(knows)[V(knows)$Interviewed=='No'])
advice.int <- delete.vertices(advice,V(knows)[V(knows)$Interviewed=='No'])
leader.int <- delete.vertices(leader,V(knows)[V(knows)$Interviewed=='No'])
infl.int <- delete.vertices(influence,V(knows)[V(knows)$Interviewed=='No'])

knows.den <- dens(knows.int)
advice.den <- dens(advice.int)
leader.den <- dens(leader.int)
infl.den <- dens(infl.int)
```

Density of knows network: `r knows.den`
Density of advice network: `r advice.den`
Density of leadership network: `r leader.den`
Density of influence network: `r infl.den`

# Leader network, interviewed people only
```{r leader-int-plot, fig.width=8}
lead.deg <- degree(leader.int,mode='in')
V(leader.int)$colour <- 'Lightblue'
V(leader.int)[lead.deg > 17]$colour <- 'Red'
plot(leader.int,layout=layout.kamada.kawai,edge.arrow.size=.8,vertex.color=V(leader.int)$colour)
```

## Dyad censuses
```{r dyads}
know.dy <- dyad.census(knows.int)
adv.dy <- dyad.census(advice.int)
lead.dy <- dyad.census(leader.int)
inf.dy <- dyad.census(infl.int)
dy.op <- matrix(cbind(know.dy$mut,know.dy$asym,know.dy$null,ecount(knows.int),knows.den,adv.dy$mut,adv.dy$asym,adv.dy$null,ecount(advice.int),advice.den,lead.dy$mut,lead.dy$asym,lead.dy$null,ecount(leader.int),leader.den,inf.dy$mut,inf.dy$asym,inf.dy$null,ecount(infl.int),infl.den),ncol=5,byrow=TRUE)
row.names(dy.op) <- c('Knows','Advice','Leader','Influence')
colnames(dy.op) <- c('Mutual','Asymmetric','Null','Total','Density')
dy.op
```
## In-degree
```{r int-centrality}
degree(leader.int,mode='in')
```
In the _leadership_ network, there is a group of five people who seem to be most prominent. This network is quite centralized, suggesting a form of hierarchical structure.

## Centralization scores, in-degree
```{r centralization}
know.cent <- centralization.degree(knows.int,mode='in')
advice.cent <- centralization.degree(advice.int,mode='in')
lead.cent <- centralization.degree(leader.int,mode='in')
infl.cent <- centralization.degree(infl.int,mode='in')
cent.op <- c(know.cent$centralization,advice.cent$centralization,lead.cent$centralization,infl.cent$centralization)
names(cent.op) <- c('Know','Advice','Leader','Influence')
cent.op
```

## Centralization scores, betweenness
```{r centralization-between}
know.cent <- centralization.betweenness(knows.int)
advice.cent <- centralization.betweenness(advice.int)
lead.cent <- centralization.betweenness(leader.int)
infl.cent <- centralization.betweenness(infl.int)
cent.op <- c(know.cent$centralization,advice.cent$centralization,lead.cent$centralization,infl.cent$centralization)
names(cent.op) <- c('Know','Advice','Leader','Influence')
cent.op
```

