Network Study
============================================
Basic network properties based on the first ten respondents (shown as the first ten in the tables, identified in red on the plots).

Ideas for analysis
-----------------------------------------------

- looking for sub-groups
- looking for gaps
- looking for stars and isolates
- looking at strength of ties, how does that compare
- looking at pattern of advice; can we identify knowledge brokers?
- what's the leadership structure?
- influential outside network? Different from leaders?
- are there bottlenecks?
- bridges or brokers? what and who?
- density? centrality?
- two-mode network?

```{r setup, echo=FALSE}
opts_chunk$set(echo=FALSE, comment='', dev='png')
```

```{r make-network}
library(igraph)
library(stringr)
library(xtable)
library(ascii)

fns <- dir(pattern='*.csv')
n <- length(fns)
nets <- list()
for (i in 1:n){
    nets[[i]] <- read.csv(fns[i],skip=5,header=FALSE,nrows=77)
}

     
el <- data.frame()
#i <- 1
for (i in 1:n){
  tmp <- nets[[i]][,c(1,2,3,6,7,8)]
  ego <- str_sub(fns[i],1,-5)
#   sp <- str_locate(ego,'[a-z][A-Z]')
#   ego <- paste(substr(ego,1,sp[1]) ,  substring(ego,sp[2]))
  sp2 <- str_locate(tmp[,1],' [A-Z]')
  alter <- paste(substr(tmp[,1],1,sp2[,1]-1),substring(tmp[,1],sp2[,2]),sep="")
#  ix <- tmp[,2] > 0
#  tmp <- tmp[ix,]

# Get rid of -1 s in advice variable
  tmp[,4] <- ifelse(tmp[,4]<=0, 0, 1)
  el <- rbind(el,data.frame(ego,alter,Knows=tmp[,2],Strength=tmp[,3],
                advice=tmp[,4],leadership=tmp[,5],influence=tmp[,6]))
}


# Reduce to non-zero edges and build a graph object
el.nonzero.edges <- subset(el, (Knows > 0 | advice > 0))


g1 <- graph.data.frame(el)

alters <- V(g1)

attribs <- read.csv('Attributes.tmp')
attribs[,1] <- attribs[,1]-1
names(attribs)[1] <- "Vertex"
# Set vertex attributes
for (i in V(g1)) {
  for (j in names(attribs)) {
    g1 <- set.vertex.attribute(g1, j, index=i, attribs[i+1,j])
  }
}

int_vertex_colors <- get.vertex.attribute(g1,"Interviewed")
colors <- c('Lightblue', 'Red')
int_vertex_colors[int_vertex_colors == 1] <- colors[1]
int_vertex_colors[int_vertex_colors == 2] <- colors[2]

```
# Summary of the complete network
```{r complete}
summary(g1)
```

# "Is this person known to you?"
## Summary of the network
```{r summary-knows}
knows <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "Knows")==0])
summary(knows)
```
## Graph of the network

```{r knows-plot, message=FALSE, fig.cap='Is person known to you', out.width='10in'}
plot(knows, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=5,
     vertex.label=NA,vertex.color=int_vertex_colors)
```
## In-degrees

```{r know-degree,comment='',results='asis'}
deg.know <- degree(knows,mode='in')
ix <- which(degree(knows)==0) - 1
knows.noiso <- delete.vertices(knows,ix)
print(xtable(data.frame(Name=alters$name,InDegree=deg.know),digits=0),type='html',html.table.attributes='border=0')
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.know), 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

# "Would you go to this person for advice?"
## Summary of the network

```{r advice}
advice <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "advice")==0])
summary(advice)
```

## Graph of the network (isolates removed)

```{r advice-plot, fig.cap='Would you go to this person for advice?'}
ix <- which(degree(advice)==0) - 1 
advice.noiso <- delete.vertices(advice,ix)
plot(advice.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
     vertex.label=NA,vertex.color=int_vertex_colors,vertex.size=3)
```
## In-degrees
```{r advice-degree, comment='', results='asis'}

deg.adv <- degree(advice,mode='in')

print(xtable(data.frame(Name=alters$name,InDegree=deg.adv),digits=0),type='html',html.table.attributes='border=0')
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.adv), 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

# "Does person occupy leadership role?"

## Summary of the network

```{r leader}
leader <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "leadership")!=1])
summary(leader)
```

## Graph of the network

```{r leader-plot, message=FALSE, fig.cap='Does the person occupy a leadership role?'}
deg.lead <- degree(leader,mode='in')

ix <- which(deg.lead==0) - 1 
leader.noiso <- delete.vertices(leader,ix)
deg.lead.no <- degree(leader.noiso, mode='in')
plot(leader.noiso,layout=layout.kamada.kawai,edge.arrow.size=1,
vertex.label=deg.lead.no,vertex.size=deg.lead.no,vertex.color=int_vertex_colors)

```
## In-degrees
```{r leader-degree, results='asis'}
lead.op <- data.frame(Name=alters$name,InDegree=deg.lead)
print(xtable(lead.op,digits=0),type='html',html.table.attributes='border=0')
#data.frame(Name=alters$name,InDegree=deg.lead)
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.lead), 1, function(X) paste(X, collapse=" | ")), sep = "\n")

```

# "Do you see the person as influential outside network?"
## Summary

```{r influence}
influence <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "influence")!=1])
summary(influence)
```

## Graph of the network (isolates removed)

```{r influence-plot, fig.cap='Is the person influential outside network?'}
deg.inf <- degree(influence, mode='in')
ix <- which(deg.inf==0) - 1 
infl.noiso <- delete.vertices(influence,ix)
deg.inf.no <- degree(infl.noiso, mode='in')
plot(infl.noiso,layout=layout.kamada.kawai,edge.arrow.size=1,vertex.label=deg.inf.no,vertex.size=deg.inf.no*2,vertex.color=int_vertex_colors)
```
## In-degree
```{r influence-degree, results='asis'}
print(xtable(data.frame(Name=alters$name,InDegree=deg.inf),digits=0),type='html',html.table.attributes='border=0')
#data.frame(Name=alters$name,InDegree=deg.inf)
#cat("Name | In degree", "--- | ---", sep="\n")
#cat(apply(data.frame(alters$name,deg.inf), 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

# Summary of all in-degrees
```{r summary, results='asis'}
op <- data.frame(Name=alters$name,Knows=deg.know,Advice=deg.adv,Leadership=deg.lead,Influence=deg.inf)
print(xtable(op,digits=0),type='html',html.table.attributes='border=0')
#cat("Name | Knows | Advice | Leadership | Influence ", "--- | --- |--- | --- | --- ", sep="\n")
#cat(apply(op, 1, function(X) paste(X, collapse=" | ")), sep = "\n")
#cat("\n")
```

# Community structure.
*Communities* are highly connected between them but with few links to other vertices.  Calculate walktrap community using the *advice* nextwork, with isolates removed. This uses a random walk algorithm to identify communities. Pascal Pons, Matthieu Latapy: Computing communities in large networks using random walks, http://arxiv.org/abs/physics/0512106. 

```{r community}
wtrp.adv <- walktrap.community(advice.noiso)
memb.adv <- community.to.membership(advice.noiso,wtrp.adv$merges,steps=12)
ix <- order(memb.adv$membership)
com.df <- data.frame(Name=get.vertex.attribute(advice.noiso,"name"),Community=memb.adv$membership)[ix,]
com.df[1:16,]
```
Modularity of this community is `r modularity(advice.noiso, memb.adv$membership)`.

## Plot of the vertices in these communities only
```{r community-plot}
vertex.colors <- c('springgreen','lightskyblue','navy','red','grey')
rem <- as.numeric(rownames(com.df[17:62,])) -1
advice.com <- delete.vertices(advice.noiso,rem)
com.vertex.colors.sub <- c(1,3,2,1,0,2,2,3,2,3,2,2,2,0,0,1)
com.vertex.colors.sub[com.vertex.colors.sub==0] <- vertex.colors[1]
com.vertex.colors.sub[com.vertex.colors.sub==1] <- vertex.colors[2]
com.vertex.colors.sub[com.vertex.colors.sub==2] <- vertex.colors[3]
com.vertex.colors.sub[com.vertex.colors.sub==3] <- vertex.colors[4]

plot(advice.com, layout=layout.kamada.kawai,edge.arrow.size=1,vertex.size=5,
       vertex.label=NA,vertex.color=com.vertex.colors.sub)
```
## Edge betweenness centrality
```{r edge-betweenness}
eb.adv <- edge.betweenness.community(advice.noiso,directed=FALSE)
eb.adv.mem <- community.to.membership(advice.noiso, eb.adv$merges, steps=9)
#modularity(advice.noiso, eb.adv.mem$membership)
#table(eb.adv.mem$membership)
#eb.adv$bridges
#eb.adv$removed.edges    
#eb.adv$edge.betweenness
brdg.colors <- rep('grey',199)
brdg.colors[eb.adv$removed.edges[1:5]+1] <- 'red'
brdg.wd <- rep(1,199)
brdg.wd[eb.adv$removed.edges[1:5]+1] <- 5
v.size <- rep(5,62)
v.size[V(advice.noiso)$name=='AnnSmallman'] <- 10

plot(advice.noiso, layout=layout.kamada.kawai,edge.arrow.size=.8,vertex.size=v.size, vertex.label=NA,edge.color=brdg.colors,edge.width=brdg.wd)
```
Plot shows the five edges that provide the most connectivity.  These are as follows:
```{r important-edges}
E(advice.noiso)[eb.adv$removed.edges[1:5]]
```
Ann Smallman looks like a particularly important bridge in the advice network.

# Network characteristics
Have to restrict to ties between respondents only.

```{r densityfunction,echo=TRUE}
dens <- function(g){
  n.edge <- ecount(g)
  n.vert <- vcount(g)
  e.max <- n.vert^2 - n.vert
  if (is.directed(g)){
    n.edge/e.max
  }
  else {
    n.edge/(e.max/2)
  }
}
```

```{r interview-networks}
knows.int <- delete.vertices(knows,10:(vcount(knows)-1))
advice.int <- delete.vertices(advice,10:(vcount(advice)-1))
leader.int <- delete.vertices(leader,10:(vcount(leader)-1))
infl.int <- delete.vertices(influence,10:(vcount(influence)-1))

knows.den <- dens(knows.int)
advice.den <- dens(advice.int)
leader.den <- dens(leader.int)
infl.den <- dens(infl.int)
```
Density of knows network: `r knows.den`
Density of advice network: `r advice.den`
Density of leadership network: `r leader.den`
Density of influence network: `r infl.den`

Summary
-------------------------------
- Some evidence of **community structure**, really need more respondents to go into this more
- Advice and info network does have some key **bridges**, but again really need more respondents as inevitable at the moment that one of the ten respondents will be in bridging position
- **Leadership** more easily measured by in-degree:
  * Four people have maximum 10: `r lead.op[lead.op[,2]==10,1]`
  * Four people with 9: `r lead.op[lead.op[,2]==9,1]`
- Structural characteristics of the network (e.g., density, centralization) need more respondents

```{r rest, echo=FALSE, eval=FALSE}

plot(degree.distribution(advice.noiso))

tkplot(advice.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
       vertex.label=NA,vertex.color=int_vertex_colors,vertex.size=deg.adv*2)
plot.coords <- tkplot.getcoords(3)
tkplot.export.postscript(3)
# Write Pajek file
write.graph(g1,'all.net',format='pajek')

deg <- degree(g1, mode='in')

plot(degree.distribution(g1))
     
bet.adv <- betweenness(advice.noiso)

x11(width=10,height=6)
plot(g1,layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.label=NA,vertex.size=deg)
dev.off()

clique.number(g1)  # largest clique has 8 members

cls <- closeness(g1, mode='in')

is.connected(g1)  # Graph is connected
clusters(g1,mode="strong")
no.clusters(g1) # 1 cluster!

cbs <- cohesive.blocks(g1)
x11(width=10,height=6)
plot.bgraph(cbs)

ebc <- edge.betweenness.community(g1)
community.to.membership(g1,ebc$merges, steps=70)

power.law.fit(deg)  # 1.44

recpr <- reciprocity(g1)  # 0.069  Very low, but prob because 


library(xtable)
xtable(data.frame(alters$name,deg.know,deg.adv,deg.lead,deg.inf),digits=0)


library(xtable)
xtable(data.frame(alters$name,deg.adv),digits=0)
#tk1 <- tkplot(g1,layout=layout.kamada.kawai,vertex.label=NA,vertex.size=5)
```