Hierarchy and homophily in a paediatric paliative care network
================================================================

Elizabeth West  
School of Health and Social Care, University of Greenwich  

David N. Barron  
Saïd Business School and Jesus College, University of Oxford  

Anne Hunt  
University of Central Lancashire  


_NB This is work in progress. Do no cite without permission_

```{r init, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
rm(list=ls())
library(knitr)
opts_chunk$set(cache=FALSE, echo=FALSE, dev='pdf')
library(igraph)
library(stringr)
library(cairoDevice)
library(knitcitations)
setwd("~/Big Study")
options(digits=2)
cite_options(tooltip = FALSE, linked = TRUE, numerical = TRUE, bibtex_data = FALSE)
bib <- read.bibtex('C:\\Users\\dbarron\\Documents\\Reputation\\Rep Measurement\\Organizations.bib')
```

Introduction
-------------------

In  contrast  to  the  hierarchical  structure  of  the  past  and  the  market  model  of  the  1980s,  the  New Labour government's reforms of the NHS were based upon collaboration and partnership working. This  model  was  first  developed  within  cancer  services.  The  Calman-Hine  Report  (1995)  had recommended that care for cancer patients should be based around managed networks of expertise and  include  the  voluntary  sector.  These  networks  that  moved  beyond  the  informal  professional groupings of the past gradually spread across a wide range of specialities. 

The  most  widely  accepted  definition  of  a  clinical  network  was  first  set  out  by  the  Scottish  Office Department  of  Health  in  1999:  “Linked  groups  of  professionals  and  organisations  from  primary, secondary   and   tertiary   care,   working   in   a   co-ordinated   manner,   unconstrained   by   existing professional [and organisational] boundaries, to ensure equitable provision of high quality, clinically effective  services.”  Networks  increasingly  came  to  be  seen  as  a  solution  to  the  co-ordination  of services for patients who require care from multi-disciplinary and multi-site teams. Partnership and network based working at all levels of care was a major policy theme in the NHS Plan (2000).  

Within  paediatric  care,  Getting  it  right  for  children  and  young  people  (2010)  proposed  that  the development of successful networks was the best means to manage the complexity of the NHS and enable children to move seamlessly from one service to another. The most effective were considered to  be  the  more  formalised  and  managed  networks  as  already  found  in  newborn,  palliative  and surgical  care. Recommendation  35  states: “There  should  be  local  networks  through  which  services can share information electronically and protocols should be agreed by the networks to provide for this.” 

As  networks  have  developed  and  their  impact  upon  the  delivery  of  health  care  become  more important,  researchers  have  looked  to  social  science  theories  to  examine  their  strengths  and weaknesses and possible implications for the NHS `r knitcitations::citep(bib[[155]])`

One important element in the success of a clinical care network, then, will be the ability of members of different organizations within the network to communicate effectively with each other and, in particular, to transfer knowledge across organizational boundaries.

Within health care, there are a wide variety of types of network, from so-called managed networks, characterised---as the name suggests---by relatively formal structures and hierarchies, to informal ``communities of practice''.  The network that we study here is closer to the latter end of the spectrum.

Literature
------------------------------
Network forms of organization have attracting an increasing amount of scholarly attention in the private as well as the public sector.  Network forms of organization are commonly viewed as a ``third way'' or organizing economic activity, an alternative to both market transactions and managerial hierarchies.  Proponents of network forms of organization argue that they are able to achieve a level of flexibility in their response to changing circumstances that is greater than hierarchies, but on the other hand they facilitate co-ordination in a way that is difficult to obtain when transactions take place in a market.

Networked businesses are often claimed to facilitate organizational learning and the diffusion of knowledge (Alvesson 2004).  There is a shift toward ``collegiality, consultation and distributed leadership'' (Ferlie et. al: p. 23), particularly in knowledge-intensive firms.

The development of managed networks in the NHS has been discussed by Ferlie et al.  They argue that there was a general trend toward the use of collaborate (as opposed to competitive) forms of governance in the public sector since the late 1990s.  The emphasis was on learning organizations and 'joined-up' policy development, ideas that required an explicit focus on communcation across professional and organizational boundaries.  Ferlie et. al show that the mnanaged network has become an important form of clinical governance in the NHS, and has had some success.  However, not all managed networks achieve their objectives.

Social and organizational boundaries have been found to be responsible for failure in organizational learning and knowledge diffusion.  For example, Ferlie et al 2005 showed that the diffusion of evidence-based clinical practice can become stuck when 

McInness et. al (2012) interviewed a range of clinical network participants in Australia, and identified a number of network characteristics that these experts believed to be crucial to network effectiveness.  One of the key characteristics they identified was the full engagement of network members across professional and organizational boundaries.  Successful networks, it was felt, were those that have developed a ''co-operative and collegial ethos'' (p. 7).

Ahgren and Axelsson argued that a critical success factor in achieving integrated health care was the development of a bottom-up (as opposed to hierarchical) approach, perhaps involving ``integrated networks'' (Ovretveit 1993).

Data and methods
---------------------------------

The members of the West Midlands Paediatric Palliative Care Network were identified through the chair  Claire  Thomas  and  a  list  of  member  names  and  e-mail  addresses  was  made  available to  the research  team  at  the  University  of  Greenwich.  An  invitation  to  participate  in  the  survey  was distributed  by  e-mail  to  network members,  who  clicked  on  the  attached  link  to  access  the  survey. The survey took approximately 15 minutes to complete. At the end of the questionnaire respondents were  asked  to  take  part  in  a  semi-structured  telephone  interview  of  about  45-60  minutes.  Any recipient who  did  not  click  the  link  in  the  e-mail  was  sent  an  electronic  reminder  after two  weeks inviting them to take part in the survey. A second reminder was sent out after four weeks to those who  had  not  yet taken  the  survey.  A paper version of  the survey was also  handed  out at  network meetings.

A total of 75 individual members of the network were identified; 22 of them completed network questionnaires. The most important aspects of the questionnaires involved people indicating which of the other 77 network members they 

1. Knew
2. Have worked with in the past 12 months
3. Would go to for advice or information about a work related problem
4. Think occupies a leadership role 
5. Is influential outside the membership network

Most of the analysis has to be restricted to the sub-networks of the 22 respondents, as only this enables us to look at a directed network, and thus to explore issues of reciprocity and mutuality in relationships.  

We look for evidence of any type of hiearchical structure in the networks of advice and leadership.  We begin with relatively simple descriptive statistics of the networks.  For example, a high proportion of mutual ties in the advice network (that is, where two people would turn to each other for advice) is an indicator of a lack of a status hierarchy.  Such a network would suggest that advice and information can flow freely through the network.

Descriptive statistics, though, suffer from it being difficult to know whether the numbers we obtain are high or low.  This can be addressed to some extent by holding constant some characteristics of the network, such as the total number of ties, and generating a large number of random networks with these properties.  We can then compare the descriptive statistics from the observed networks with those from the random networks to get a sense of how much they differ from those that would have been expected if network ties were created purely at random.

Finally, we would also like to explore more models that allow us to investigate whether the existence of ties in these networks can be explained by reference to characteristics of individual network members.  For example, it is possible that longer-serving members are more likely to be viewed as leaders or approached for advice because they are viewed as having greater expertise.  It is also possible that there are 'homophily' effects in tie formation.  This well-known characteristic of many networks is often summarised with the saying, 'Birds of a feather flock together'.  In these networks, we are particularly interested to see whether professional background helps explain the existence of ties; if it does, then this suggests that professional boundaries could present barriers to effective information flow within the network.  This type of study involves the use of exponential random graph models (ergm).

Results
-------------------

### Advice and information

```{r readdata, cache=FALSE}
fns <- dir(pattern='*.csv')  # Gets all csv file names
# There is a file Professions.csv that has to be excluded
ix2 <- which(str_detect(fns,"Profession"))
fns <- fns[-ix2]
ix <- str_detect(fns,"Orgs")  # TRUE when filename has Orgs in it, FALSE otherwise
org.fns <- fns[ix]  # Vector of organizational data files
fns <- fns[!ix]  # Vector of interpersonal data filenames
n <- length(fns)  # number of interperson filenames
nets <- list()   # nets will hold data
for (i in 1:n){
    nets[[i]] <- read.csv(fns[i],skip=5,header=FALSE,nrows=77)
}
```

```{r personalnets, cache=FALSE}
el <- data.frame()  # Data will be in this dataframe

for (i in 1:n){
  tmp <- nets[[i]][,c(1,2,3,4,6,7,8)]  # don't include col 5 as this is attribute
  ego <- str_sub(fns[i],1,-5)  # Take ego name from the original filename
  sp2 <- str_locate(tmp[,1],' [A-Z]')  # Locate space in alter's name
  # Then use this location to create alter name without space
  alter <- paste(substr(tmp[,1],1,sp2[,1]-1),substring(tmp[,1],sp2[,2]),sep="")
  # Correct a couple of inconsistencies in people's names
  alter <- ifelse(alter=='RachaelWiliams','RachaelWilliams',alter)  
  alter <- ifelse(alter=='SuEdwards','SueEdwards',alter)
  # Get rid of -1 s in advice variable
  tmp[,5] <- ifelse(tmp[,5]<=0, 0, 1)
  tmp[,6] <- ifelse(tmp[,6]<=0, 0, 1)
  tmp[,7] <- ifelse(tmp[,7]<=0, 0, 1)
  # Create data frame by stacking data from each ego
  el <- rbind(el,data.frame(ego,alter,Knows=tmp[,2],Strength=tmp[,3],worked=tmp[,4,],
                advice=tmp[,5],leadership=tmp[,6],influence=tmp[,7]))
}
ix <- order(el$ego)
el <- el[ix,]

```

```{r attributes}
## Basic attribute data, has whether interviewed and org name
atts <- read.table("Attributes.txt",header=TRUE,stringsAsFactors=FALSE)
atts <- atts[,-3]
orgs <- sort(unique(atts[,3]))
org.ids <- 1:length(orgs)
orgs.dta <- data.frame(OrganizationName=orgs, OrganizationID = org.ids)

atts.2 <- merge(atts,orgs.dta,by="OrganizationName")
  ix2 <- order(atts.2$Name)
atts.2 <- atts.2[ix2,]

## > names(atts)
##[1] "Name"             "Interviewed"      "OrganizationName"
 
## Data on professional background
prof <- read.csv('Profession.csv',stringsAsFactor=FALSE)

### Treat Health Vistor as Nurse (there is only one HV)
ix.hv <- prof$Prof.Name == "Health Visitor"
prof$Prof.Name[ix.hv] <- "Nurse"
prof$Profession[ix.hv] <- 1
prof$Name <- paste(prof[,1],prof[,2],sep='')
prof <- prof[,-c(1,2)]
prof$Profession <- factor(prof$Profession,levels=c(1:6,99),labels=c('Nurse','Doctor','Allied HP','Social service','Manager','Other','Unknown'))
prof$Prof.Name <- factor(prof$Prof.Name)
ix.na <- prof$Prof.Name == 'Unknown'
prof[ix.na,"Prof.Name"] <- NA
prof[ix.na,"Profession"] <- NA
all.atts <- merge(atts.2,prof,by='Name',all.x=TRUE)
```

```{r mergeorgs}
## Test 
ix.e <- order(el$ego)
el <- el[ix.e,]

if (all.equal(atts.2[atts.2$'Interviewed'=='Yes','Name'],levels(el$ego))){

  # Merge network data and attribute data
  el.2 <- merge(el,all.atts,by.x="ego",by.y='Name',all.x=TRUE)
  names(el.2)[9] <- 'EgoOrganization'
  names(el.2)[11] <- 'EgoOrganizationID'
  names(el.2)[12] <- 'EgoProfession'
  names(el.2)[13] <- 'EgoProf.Name'
  
  el.3 <- merge(el.2,all.atts,by.x="alter",by.y='Name',all=TRUE)
  names(el.3)[15] <- 'AlterOrganization'
  names(el.3)[17] <- 'AlterOrganizationID'
  names(el.3)[18] <- 'AlterProfession'
  names(el.3)[19] <- 'AlterProf.Name'
  
  ix.e2 <- order(el.3$ego)
  el.3 <- el.3[ix.e2,]
  
    
  el.org <- el.3[,c(9,13,3,5,6,7,8)]
  attrib.org <- el.3[,c(10,2,1,4,8,12,9)]
  names(attrib.org)[7] <- 'Interviewed'
}
```

```{r graps}
# Remove zero edges
#el.nonzero.edges <- subset(el, (Knows > 0 | advice > 0))

# Create igraph object
g1 <- graph.data.frame(el)
alters <- V(g1)


### Attach vertex attributes
for (i in V(g1)) {
  for (j in names(all.atts)) {
    ix <- match(V(g1)$name[i],all.atts[,1])
    if (is.na(ix)) stop("Invalid name")
    g1 <- igraph::set.vertex.attribute(g1, j, index=i, all.atts[ix,j])
  }
}

## Setup vertex colors for plotting
int_vertex_colors <- igraph::get.vertex.attribute(g1,"Interviewed.x")
colors <- c('Lightblue', 'Red')
int_vertex_colors[int_vertex_colors == 'No'] <- colors[1]
int_vertex_colors[int_vertex_colors == 'Yes'] <- colors[2]

prof_vertex_colors <- as.numeric(igraph::get.vertex.attribute(g1,'Profession'))
colors <- c('Lightblue', 'Red','Green','Yellow','Orange','Purple')
prof_vertex_colors[prof_vertex_colors == 1] <- colors[1]
prof_vertex_colors[prof_vertex_colors == 2] <- colors[2]
prof_vertex_colors[prof_vertex_colors == 3] <- colors[3]
prof_vertex_colors[prof_vertex_colors == 4] <- colors[4]
prof_vertex_colors[prof_vertex_colors == 5] <- colors[5]
prof_vertex_colors[prof_vertex_colors == 6| is.na(prof_vertex_colors)] <- colors[6]

```


```{r grc}
# Only implemented for unweighted, directed graphs
grc <- function(g, local=FALSE){
  n <- length(V(g))
  local.rc <- vector(length=n)
  for (i in 1:n){
    local.rc[i] <- (length(subcomponent(g, i, mode='out')) - 1) / (n-1)
  }
  Cmax <- max(local.rc)
  glob.reach <- sum(Cmax - local.rc)/(n-1)
  if (local) {
    return(list(grc=glob.reach,local=local.rc))
  }
  else {
    return(glob.reach)
  }
}
```

```{r separatenets}
advice <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "advice")==0])
advice.noiso <- delete.vertices(advice,V(advice)[degree(advice)==0])
leader <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "leadership")!=1])
leader.noiso <- delete.vertices(leader,V(leader)[degree(leader)==0])
advice.int <- delete.vertices(advice,V(advice)$Interviewed.x=='No')
leader.int <- delete.vertices(leader,V(leader)$Interviewed.x=='No')
deg.adv <- degree(advice.noiso,mode='in')
influence <- delete.edges(g1,E(g1)[get.edge.attribute(g1,name='influence')==0])
infl.noiso <- delete.vertices(influence,V(influence)[degree(influence)==0])
infl.int <- delete.vertices(influence,V(influence)$Interviewed.x=='No')
deg.inf <- degree(infl.noiso, mode='in')
```

```{r advplots, fig.width=4, fig.height=4 }
plot(advice.noiso, layout=layout.kamada.kawai,edge.arrow.size=.8,
     vertex.label=NA,vertex.color=prof_vertex_colors,vertex.size=deg.adv, main="Advice network")
legend('topleft',fill=c('Lightblue', 'Red','Green','Yellow','Orange','Purple'), legend=c('Nurse','Doctor','Allied HP','Social services','Manager','Other/Unknown'))
```

The advice network is plotted, with the seven people who received no nominations omitted.  The size of the circles is proportional to the number of times the person was nominated as a source of advice.  The degree distribution is shown below.

```{r adv.deg}

deg.dist <- function (graph, cumulative = FALSE, ...) 
{
    if (!igraph::is.igraph(graph)) {
        stop("Not a graph object")
    }
    cs <- igraph::degree(graph, ...)
    hi <- hist(cs, -1:max(cs), plot = FALSE)$density
    if (!cumulative) {
        res <- hi
    }
    else {
        res <- rev(cumsum(rev(hi)))
    }
    res
}
plot(deg.dist(advice, mode='in'), main='Advice in-degree distribution',ylab='',type='h',xlab='In-degree') 
```

As is typical, there are a relatively small number of people who have a high degree.  Bearing in mind that the maximum possible degree is 22 (as only 22 people participated in the study), it is notable that seven people receive requests for advice from 16 or more people, which is equal to the number of people who are not asked for advice by anyone.  This is in itself evidence of a sort of status hierarchy in the network, although we cannot know whether those people who give a lot of advice also frequently seek advice.  If they do, then although that would suggest they are in some sense prominent in the network, it would give rather a different picture from one in which there are people who are seen as experts and who do not themselves need to seek advice.  However, we can investigate such questions in the sub-network of respondents.

```{r subadvplot}
adv.deg <- degree(advice.int, mode='in')
plot(advice.int, edge.arrow.size=.8, vertex.color=prof_vertex_colors, layout=layout.kamada.kawai, vertex.size=adv.deg, vertex.label=NA,main='Advice')
legend('topleft',fill=c('Lightblue', 'Red','Green','Yellow','Orange','Purple'), legend=c('Nurse','Doctor','Allied HP','Social services','Manager','Other/Unknown'))

dc.adv.int <- igraph::dyad.census(advice.int)
```

In this sub-network we can again look for evidence of hierarchy.  The global reach centrality for this subnetwork is zero, which implies no hiearchy at all.  This is because all members of the network are strongly connected to each other.  In a traditional hierarchy, where the most senior manager has authority over his or her deputies, who in turn have authority over their subordinates, and so on, are not strongly connected because there are no paths linking subordinates to people above them in the hierarchy. The advice and information network here does not have this quality as some pairs of people ask each other for advice.  This reciprocity of advice-giving means that there are directed paths between all members of the network, and hence no hierarchy in the formal sense.

```{r randnets}
dyad.rand <- function(g, niter=100, rwiter=100){
  indeg <- degree(g, mode='in')
  outdeg <- degree(g, mode='out')
  rand.dc.mut <- vector(length=niter)
  rand.dc.asym <- vector(length=niter)
  rand.dc.null <- vector(length=niter)
  for (i in 1:niter){
    tmp <- degree.sequence.game(outdeg, indeg, method='simple')
    tmp.rand <- rewire(tmp, niter=rwiter)
    rand.dc <- dyad.census(tmp.rand)
    rand.dc.mut[i] <- rand.dc$mut
    rand.dc.asym[i] <- rand.dc$asym
    rand.dc.null[i] <- rand.dc$null
  }
  list(Mutual=rand.dc.mut,Asym=rand.dc.asym,Null=rand.dc.null)
}
```

```{r randdyad, cache=FALSE}
set.seed(1234)
dc.r <- dyad.rand(advice.int,1000)
dc.means <- sapply(dc.r,mean)
dc.sd <- sapply(dc.r,sd)
```

There may nevertheless still be hierarchies in a less formal sense: status hierachies, for example, might exist if some people are more likely to be asked for advice than others, as indeed is the case (see figure x).  We can get some sense of the degree of reciprocity by looking at the number of "mutual" dyads.  In the sub-network, the number of mutual dyads is ```r dc.adv.int$mut``` while the number of asymmetric dyads is ```r dc.adv.int$asym``` and the number of dyads with no ties in either direction is ```r dc.adv.int$null```.  

To get a sense of the interpretation of these numbers, I generated 1000 random networks with  the same in and out degree as the advice sub-network using the `degree.sequence.game` function in `R`, and then randomly rewired 100 of the ties in each of these networks.  The mean number of mutual, asymmetric and null dyads in these 1000 networks was ```r dc.means[1]``` (standard deviation = ```r dc.sd[1]```), ```r dc.means[2]``` (```r dc.sd[2]```), and ```r dc.means[3]``` (```r dc.sd[3]```), respectively.  Notice that there are more than twice as many mutual ties in the observed advice and information sub-network than in the random networks, which  suggests that there is a high degree of sharing of information and expertise in this network, implying a very non-hierarchical network structure.  On the other hand, there are also twice as many pairs with no connection at all in the observed network than in the random network.  One possible explanation for this pattern is that there is some sort of assortative matching, or homophily, taking place; for example, people from the same professional background may be more likely to seek each other out as sources of advice and information.  This would be a potential problem for the network if true, as it would be a barrier to a source of learning and integration across professional boundaries.

### Exponential random graph model


```{r proftable, results='asis'}
library(xtable)
print(xtable(table(prof$Profession)), type='HTML')
```
Professional background of network members

To evaluate any propensity for netowrk members to tend to approach members of their own profession for advice, we need to control for the numbers of each profession in the network. In the table above we can see that there are more than twice as many nurses as there are doctors.  Therefore, even if nurses were choosing people to turn to for advice at random, they would be twice as likely to choose a nurse as a doctor.


```{r ergm1, message=FALSE, warning=FALSE, results='hide'}
library(network)
library(ergm)
detach(package:igraph)
adv.adj <- igraph::get.adjacency(advice, sparse=FALSE)
adv.int.adj <- igraph::get.adjacency(advice.int, sparse=FALSE)
adv.inter <- igraph::V(advice)$Interviewed.x
adv.net <- network::network(adv.adj)
adv.int.net <- network::network(adv.int.adj)
network::set.vertex.attribute(adv.net,'Interviewed',adv.inter)

adv.prof <- igraph::V(advice)$Profession
network::set.vertex.attribute(adv.net,'Profession',adv.prof)
adv.int.prof <- igraph::V(advice.int)$Profession
network::set.vertex.attribute(adv.int.net,'Profession',adv.int.prof)

#e0 <- ergm(adv.net ~ edges + odegree(0))
#summary(e0)

e1 <- ergm(adv.net ~  edges  + odegree(0) + idegree(0) + mutual + nodeifactor('Profession',base=c(3,4,6)) + nodematch('Profession', diff=TRUE, keep=c(1,2,5))  ,control=control.ergm(seed=10, MCMLE.maxit=50))

#e1.int <- ergm(adv.int.net ~  edges + nodeifactor('Profession',base=c(3,4,6)) + nodematch('Profession', diff=TRUE, keep=c(1,2)) + gwidegree(1) ,control=control.ergm(seed=10, MCMLE.maxit=50))

mix <- mixingmatrix(adv.net,'Profession')
mix.mat <- mix$matrix

b.e1 <- coef(e1)
se.e1 <- sqrt(diag(summary(e1)$asycov))
#e1.gof <- gof(e1)
#plot(e1.gof)

```
The table shows the crosstabulation of ego's and alter's profession in the advice network.

```{r mixmat1, results='asis'}
rownames(mix.mat) <- c('Nurse','Doctor','Allied HP','Social service','Manager','Other')
colnames(mix.mat) <- rownames(mix.mat)
print(xtable(mix.mat, digits=0), type='HTML')
```

One way to  investigate whether there is evidence of assortative matching controling for the frequency of different professional groups in the network is by means of an exponential random graph model (ERGM). These models provide a method for the statistical analysis of social networks.   In these models, the probability of a particular, observed network, $X=x$ conditional on a vector of parameters, $\theta$, nodes is given by:

$$
  \Pr (X=x | \theta) = \frac{1}{\kappa(\theta)} \exp\{ \theta_1 z_1(x) + \theta_2 z_2(x) + \cdots + \theta_p z_p(x) \}.
$$

The functions $z_k(x)$ are counts of configurations in the network $x$, the parameters, $\theta_k$ weight the relative importance of the corresponding network configuration, and the function $\kappa(\theta)$ is a normalising term `r knitcitations::citep(bib[[156]])`.  An intuition behind this method is that we compare the 

>"observed network to the other possible ways it could be arranged.  In all networks
there are a finite number of ways that ties can be arranged....Then, for a distribution of [networks], we have to assign a probability to each [network].  Model estimation assigns probabilities so that an observed network is central, not extreme, in terms of the effects being modelled" `r knitcitations::citep(bib[[156]])`.

The table below shows the result sof an estimate that includes main effects of professional background as well as variables that measure the propensity of people in the three main professional groups to select members of their own profession.


                 | Estimate    | Standard error
-----------------|-------------|------------------------
Edges            |`r b.e1[1]`* | `r se.e1[1]`
Zero outdegree   |`r b.e1[2]`* | `r se.e1[2]`
Zero indegree    |`r b.e1[3]`* | `r se.e1[3]`
Mutual ties      |`r b.e1[4]`* | `r se.e1[4]`
Nurses           |`r b.e1[5]`* | `r se.e1[5]`
Doctors          |`r b.e1[6]`* | `r se.e1[6]`
Managers         |`r b.e1[7]`  | `r se.e1[7]`
Nurse homophily  | `r b.e1[8]` | `r se.e1[8]`
Doctor homophily |`r b.e1[9]`* | `r se.e1[9]`
Manager homophily| `r b.e1[10]` | `r se.e1[10]`  
AIC              |`r AIC(e1)`  |
Note: * = *p* < .05

The model includes a term for the number of ties (or edges) in the network, a term that is usually said to serve the same function as the intercept in linear regression, representing the baseline propensity for ties to form (that is, for people to name others to whom they turn for advice).  This network is, of course, unusual in that there are a large number of nodes with an outdegree of zero---by definition, people who did not participate in the study cannot have nominated any advisors.  Therefore, an additional control variable for nodes with a zero outdegree is also included in the model.  As we have already noted a tendency for mutual ties to be more common than expected, we also include a term to control for this.  We also include main effects to control for the tendancy of nurses, doctors and managers to name advisors; the fact that these are all positive and statistically significant implies that members of these groups have more advisors, on average, than members of the other professions (allied health professions and social service professions).

We can see that there remains a statitically significant tendancy for doctors to prefer other doctors as advisors.  The estimated homophily effect for nurses is negative, implying that nurses are less likely to choose other nurses as advisors than would be expected by chance, but this estimate is not statistically significant.  Similarly, the postitive estimate for managers suggests that there is some tendancy for managers to disproportionately approach other managers for advice, but again this effect is not statistically significant.

### Leadership

```{r leadplots }
detach(package:ergm)
detach(package:network)

library(igraph)
deg.lead <- degree(leader.noiso,mode='in')

plot(leader.noiso, layout=layout.kamada.kawai,edge.arrow.size=.8,
     vertex.label=NA,vertex.color=prof_vertex_colors,vertex.size=deg.lead, main="Leadership network")
legend('topleft',fill=c('Lightblue', 'Red','Green','Yellow','Orange','Purple'), legend=c('Nurse','Doctor','Allied HP','Social services','Manager','Other/Unknown'))

dc.lead.int <- igraph::dyad.census(leader.int)

```

The participants in the survey were also asked to nominate those people that they though occupied leadership roles within the network.  The network is displayed in the figure, with the people who received no nominations omitted.

Generally, we would expect to see more evidence of a hierarchical structure in a leadership network.  In traditional organizational structures, leadership relationships cannot be reciprocal by definition.  However, the situation may be different in this network, lacking as it does any _formal_ authority structure.  Survey respondents were asked to identify people who occupied leadership roles in the nextwork, and so it is possible that these ties could be reciprocal as there could be a number of people that are considered to act as leaders in some sense or another.

Indeed, when the network diagram  aboove shows that a significant number of people in the network were identified as leaders.  In fact ```r sum(degree(leader, mode='in')>0) ``` people were nominated as leaders by at least one colleague out of the 78 people in the network, a remarkably high proportion.

Turning as before to the subnetwork of those peoplel who were interviewed, the global reach centrality score is ```r grc(leader.int)```, slightly more than the advice network, but still very low, indicating a low level of hiearchy.  In the sub-network, the number of mutual dyads is ```r dc.lead.int$mut``` while the number of asymmetric dyads is ```r dc.lead.int$asym``` and the number of dyads with no ties in either direction is ```r dc.lead.int$null```. The leadership network is, then, somewhat more hierarchical than the advice network, but for a network on this type of relation there is in fact a very high degree of mutuality in this network, much more than one would expect in a random network.


```{r leaddeg}
plot(deg.dist(leader, mode='in'), main='Leadership in-degree distribution',ylab='',type='h',xlab='In-degree') 
```

```{r ergmlead, message=FALSE, warning=FALSE, results='hide'}
library(network)
library(ergm)
detach(package:igraph)
lead.adj <- igraph::get.adjacency(leader, sparse=FALSE)
lead.int.adj <- igraph::get.adjacency(leader.int, sparse=FALSE)
leader.inter <- igraph::V(leader)$Interviewed.x
lead.net <- network::network(lead.adj)
lead.int.net <- network::network(lead.int.adj)
network::set.vertex.attribute(lead.net,'Interviewed',leader.inter)

lead.prof <- igraph::V(leader)$Profession
network::set.vertex.attribute(lead.net,'Profession',lead.prof)
lead.int.prof <- igraph::V(leader.int)$Profession
network::set.vertex.attribute(lead.int.net,'Profession',lead.int.prof)

#e0 <- ergm(adv.net ~ edges + odegree(0))
#summary(e0)

e2 <- ergm(lead.net ~  edges  + odegree(0) + idegree(0) + mutual + nodeifactor('Profession',base=c(3,4,6)) + nodematch('Profession', diff=TRUE, keep=c(1,2,5))  ,control=control.ergm(seed=10, MCMLE.maxit=50))

#e1.int <- ergm(adv.int.net ~  edges + nodeifactor('Profession',base=c(3,4,6)) + nodematch('Profession', diff=TRUE, keep=c(1,2)) + gwidegree(1) ,control=control.ergm(seed=10, MCMLE.maxit=50))

b.e2 <- coef(e2)
se.e2 <- sqrt(diag(summary(e2)$asycov))
#e1.gof <- gof(e1)
#plot(e1.gof)

```

We can again explore the extent to which mutuality is a reflection of people tending to nominate members of their own profession as leaders, using a similar model to that used for the advice network above.  The results are shown below:

                 | Estimate    | Standard error
-----------------|-------------|------------------------
Edges            |`r b.e2[1]`* | `r se.e2[1]`
Zero outdegree   |`r b.e2[2]`* | `r se.e2[2]`
Zero indegree    |`r b.e2[3]`* | `r se.e2[3]`
Mutual ties      |`r b.e2[4]`* | `r se.e2[4]`
Nurses           |`r b.e2[5]`  | `r se.e2[5]`
Doctors          |`r b.e2[6]`* | `r se.e2[6]`
Managers         |`r b.e2[7]`* | `r se.e2[7]`
Nurse homophily  |`r b.e2[8]`  | `r se.e2[8]`
Doctor homophily |`r b.e2[9]`  | `r se.e2[9]`
Manager homophily|`r b.e2[10]` | `r se.e2[10]`  
AIC              |`r AIC(e2)`  |
Note: * = *p* < .05

One possible interpretation of this is that the network is a genuine alternative to a hierarchical structure, and as such is a very different model from that of the managed networks that are becoming a more prominent feature of the NHS.  


### Influence

```{r infplots }
detach(package:ergm)
detach(package:network)

library(igraph)
inf.deg <- degree(infl.noiso,mode='in')

plot(infl.noiso, layout=layout.kamada.kawai,edge.arrow.size=.8,
     vertex.label=NA,vertex.color=prof_vertex_colors,vertex.size=inf.deg, main="Influence network")
legend('topleft',fill=c('Lightblue', 'Red','Green','Yellow','Orange','Purple'), legend=c('Nurse','Doctor','Allied HP','Social services','Manager','Other/Unknown'))

dc.infl.int <- igraph::dyad.census(infl.int)
```

Turning as before to the subnetwork of those peoplel who were interviewed, the global reach centrality score is ```r grc(infl.int)```, slightly more than the advice network, but still very low, indicating a low level of hiearchy.  In the sub-network, the number of mutual dyads is ```r dc.infl.int$mut``` while the number of asymmetric dyads is ```r dc.infl.int$asym``` and the number of dyads with no ties in either direction is ```r dc.infl.int$null```. The influence network is, then, somewhat more hierarchical than the advice network, but for a network on this type of relation there is in fact a very high degree of mutuality in this network, much more than one would expect in a random network.


```{r ergminf, message=FALSE, warning=FALSE, results='hide'}
library(network)
library(ergm)
detach(package:igraph)
infl.adj <- igraph::get.adjacency(influence, sparse=FALSE)
#lead.int.adj <- igraph::get.adjacency(leader.int, sparse=FALSE)
#leader.inter <- igraph::V(leader)$Interviewed.x
infl.net <- network::network(infl.adj)
#lead.int.net <- network::network(lead.int.adj)
#network::set.vertex.attribute(lead.net,'Interviewed',leader.inter)

infl.prof <- igraph::V(influence)$Profession
network::set.vertex.attribute(infl.net,'Profession',infl.prof)
#lead.int.prof <- igraph::V(leader.int)$Profession
#network::set.vertex.attribute(lead.int.net,'Profession',lead.int.prof)

#e0 <- ergm(adv.net ~ edges + odegree(0))
#summary(e0)

e3 <- ergm(infl.net ~  edges  + odegree(0) + idegree(0) + mutual + nodeifactor('Profession',base=c(3,4,6)) + nodematch('Profession', diff=TRUE, keep=c(1,2,5))  ,control=control.ergm(seed=10, MCMLE.maxit=50))

#e1.int <- ergm(adv.int.net ~  edges + nodeifactor('Profession',base=c(3,4,6)) + nodematch('Profession', diff=TRUE, keep=c(1,2)) + gwidegree(1) ,control=control.ergm(seed=10, MCMLE.maxit=50))

b.e3 <- coef(e3)
se.e3 <- sqrt(diag(summary(e3)$asycov))
#e1.gof <- gof(e1)
#plot(e1.gof)

```

                 | Estimate    | Standard error
-----------------|-------------|------------------------
Edges            |`r b.e3[1]`* | `r se.e3[1]`
Zero outdegree   |`r b.e3[2]`* | `r se.e3[2]`
Zero indegree    |`r b.e3[3]`* | `r se.e2[3]`
Mutual ties      |`r b.e3[4]`* | `r se.e3[4]`
Nurses           |`r b.e3[5]`  | `r se.e3[5]`
Doctors          |`r b.e3[6]`* | `r se.e3[6]`
Managers         |`r b.e3[7]`  | `r se.e3[7]`
Nurse homophily  |`r b.e3[8]`* | `r se.e3[8]`
Doctor homophily |`r b.e3[9]`  | `r se.e3[9]`
Manager homophily|`r b.e3[10]` | `r se.e3[10]`
AIC              |`r AIC(e3)`  |
Note: * = *p* < .05

# References

```{r refs, results='asis'}
bibliography()
```

# Appendix

## Goodness of fit plots

### Advice network

```{r advicegof, message=FALSE, cache=FALSE, eval=FALSE}
library(network)
library(ergm)
detach(package:igraph)
gof.advice <- gof(e1)
plot(e1)
plot(gof.advice)
```

### Leadership network

```{r leadgof, message=FALSE, cache=FALSE, eval=FALSE}
library(network)
library(ergm)
gof.lead <- gof(e2)
plot(e2)
plot(gof.lead)
```

### Influence network

```{r infgof, message=FALSE, cache=FALSE, eval=FALSE}
gof.inf <- gof(e3)
plot(e3)
plot(gof.inf)
```
