Network analysis using Big Study data
========================================================

```{r init, echo=FALSE, message=FALSE}
opts_chunk$set(cache=FALSE, echo=FALSE)
library(igraph)
library(stringr)

setwd("~/Big Study")
```
# Data preparation 

## Read interpersonal network data

Each person interviewed has two csv files, one has the name FirstnameLastname.csv,
the other the name FirstnameLastnameOrgs.csv.  The first has data on relationships with other invividuals, the second relationships with other organizations.  
**DO NOT PUT OTHER csv FILES IN THE DIRECTORY**

```{r readdata, cache=TRUE}
fns <- dir(pattern='*.csv')  # Gets all csv file names
# There s a file Professions.csv that has to be excluded
ix2 <- which(str_detect(fns,"Profession"))
fns <- fns[-ix2]
ix <- str_detect(fns,"Orgs")  # TRUE when filename has Orgs in it, FALSE otherwise
org.fns <- fns[ix]  # Vector of organizational data files
fns <- fns[!ix]  # Vector of interpersonal data filenames
n <- length(fns)  # number of interperson filenames
nets <- list()   # nets will hold data
for (i in 1:n){
    nets[[i]] <- read.csv(fns[i],skip=5,header=FALSE,nrows=77)
}
```

## Data is taken from each component of the list and combined.

```{r personalnets, cache=TRUE}
el <- data.frame()  # Data will be in this dataframe

for (i in 1:n){
  tmp <- nets[[i]][,c(1,2,3,4,6,7,8)]  # don't include col 5 as this is attribute
  ego <- str_sub(fns[i],1,-5)  # Take ego name from the original filename
  sp2 <- str_locate(tmp[,1],' [A-Z]')  # Locate space in alter's name
  # Then use this location to create alter name without space
  alter <- paste(substr(tmp[,1],1,sp2[,1]-1),substring(tmp[,1],sp2[,2]),sep="")
  # Correct a couple of inconsistencies in people's names
  alter <- ifelse(alter=='RachaelWiliams','RachaelWilliams',alter)  
  alter <- ifelse(alter=='SuEdwards','SueEdwards',alter)
  # Get rid of -1 s in advice variable
  tmp[,4] <- ifelse(tmp[,4]<=0, 0, 1)
  # Create data frame by stacking data from each ego
  el <- rbind(el,data.frame(ego,alter,Knows=tmp[,2],Strength=tmp[,3],
                advice=tmp[,4],leadership=tmp[,5],influence=tmp[,6]))
}
ix <- order(el$ego)
el <- el[ix,]

```

## Add attribute data

```{r attributes}
## Basic attribute data, has whether interviewed and org name
atts <- read.table("Attributes.txt",header=TRUE,stringsAsFactors=FALSE)
atts <- atts[,-3]
orgs <- sort(unique(atts[,3]))
org.ids <- 1:length(orgs)
orgs.dta <- data.frame(OrganizationName=orgs, OrganizationID = org.ids)

atts.2 <- merge(atts,orgs.dta,by="OrganizationName")
  ix2 <- order(atts.2$Name)
atts.2 <- atts.2[ix2,]

## > names(atts)
##[1] "Name"             "Interviewed"      "OrganizationName"
 
## Data on professional background
prof <- read.csv('Profession.csv',stringsAsFactor=FALSE)
prof$Name <- paste(prof[,1],prof[,2],sep='')
prof <- prof[,-c(1,2)]
prof[,1] <- factor(prof[,1],levels=c(1:6,99),labels=c('Nurse','Doctor','Allied HP','Social service','Manager','Other','Unknown'))
ix <- prof[,1] == 'Unknown'
prof[ix,1] <- NA
all.atts <- merge(atts.2,prof,by='Name',all.x=TRUE)
all.atts <- all.atts[,-7]
names(all.atts)
names(all.atts)[3] <- 'Interviewed'
# [1] "Name"             "Interviewed"      "OrganizationName" "Profession"      
```

```{r mergeorgs}
## Test 
ix.e <- order(el$ego)
el <- el[ix.e,]

if (all.equal(atts.2[atts.2$'Interviewed'=='Yes','Name'],levels(el$ego))){

  # Merge network data and attribute data
  el.2 <- merge(el,all.atts,by.x="ego",by.y='Name',all.x=TRUE)
  names(el.2)[8] <- 'EgoOrganization'
  names(el.2)[10] <- 'EgoOrganizationID'

  el.3 <- merge(el.2,all.atts,by.x="alter",by.y='Name',all=TRUE)
  names(el.3)[12] <- 'AlterOrganization'
  names(el.3)[14] <- 'AlterOrganizationID'
  ix.e2 <- order(el.3$ego)
  el.3 <- el.3[ix.e2,]
  names(el.3)
#  [1] "alter"               "ego"                 "Knows"              
#  [4] "Strength"            "advice"              "leadership"         
#  [7] "influence"           "EgoOrganization"     "Interviewed.x"      
# [10] "EgoOrganizationID"   "Profession.x"        "AlterOrganization"  
# [13] "Interviewed.y"       "AlterOrganizationID" "Profession.y"   
  
  el.org <- el.3[,c(8,12,3,5,6,7)]
  attrib.org <- el.3[,c(10,2,1,4,8,12,9)]
  names(attrib.org)[7] <- 'Interviewed'
}
```
## Create `igraph` objects

```{r graps}
# Remove zero edges
el.nonzero.edges <- subset(el, (Knows > 0 | advice > 0))

# Create igraph object
g1 <- graph.data.frame(el)
summary(g1)
alters <- V(g1)


### Attach vertex attributes
for (i in V(g1)) {
  for (j in names(all.atts)) {
    ix <- match(V(g1)$name[i],all.atts[,1])
    if (is.na(ix)) stop("Invalid name")
    g1 <- igraph::set.vertex.attribute(g1, j, index=i, all.atts[ix,j])
  }
}

## Setup vertex colors for plotting
int_vertex_colors <- igraph::get.vertex.attribute(g1,"Interviewed")
colors <- c('Lightblue', 'Red')
int_vertex_colors[int_vertex_colors == 'No'] <- colors[1]
int_vertex_colors[int_vertex_colors == 'Yes'] <- colors[2]

prof_vertex_colors <- igraph::get.vertex.attribute(g1,'Profession')
colors <- c('Lightblue', 'Red','Green','Yellow','Orange')
prof_vertex_colors[prof_vertex_colors == 'Nurse'] <- colors[1]
prof_vertex_colors[prof_vertex_colors == 'Doctor'] <- colors[2]
prof_vertex_colors[prof_vertex_colors == 'Allied HP'] <- colors[2]
prof_vertex_colors[prof_vertex_colors == 'Social service'] <- colors[2]
prof_vertex_colors[prof_vertex_colors == 'Manager'] <- colors[2]

```


```{r grc}
# Only implemented for unweighted, directed graphs
grc <- function(g, local=FALSE){
  n <- length(V(g))
  local.rc <- vector(length=n)
  for (i in 1:n){
    local.rc[i] <- (length(subcomponent(g, i, mode='out')) - 1) / (n-1)
  }
  Cmax <- max(local.rc)
  glob.reach <- sum(Cmax - local.rc)/(n-1)
  if (local) {
    return(list(grc=glob.reach,local=local.rc))
  }
  else {
    return(glob.reach)
  }
}
```


# Initial analysis of 'Known' network

```{r knows}
# Create separate network
knows <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "Knows")==0])
deg.know <- degree(knows,mode='in')
deg.know.tab <- table(deg.know)
# Create network with isolates removed
knows.noiso <- delete.vertices(knows,V(knows)[degree(knows)==0])

plot(knows.noiso, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=5,
     vertex.label=NA,vertex.color=int_vertex_colors)

plot(knows.noiso, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=5,
     vertex.label=NA,vertex.color=prof_vertex_colors)

#plot(degree.distribution(knows))
```


## Advice network

```{r advice}
advice <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "advice")==0])
summary(advice)
deg.adv <- degree(advice,mode='in')

advice.noiso <- delete.vertices(advice,V(advice)[degree(advice)==0])
deg.adv <- degree(advice,mode='in')
#plot(degree.distribution(advice.noiso))

plot(advice, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
     vertex.label=NA,vertex.color=int_vertex_colors,vertex.size=3)
plot(advice.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
     vertex.label=NA,vertex.color=prof_vertex_colors,vertex.size=3)

#plot(degree.distribution(advice))
```

```{r advBetween, eval=FALSE}
### Betweenness community structure


eb.adv <- edge.betweenness.community(advice.noiso,directed=FALSE)
eb.adv.mem <- community.to.membership(advice.noiso, eb.adv$merges, steps=9)
modularity(advice.noiso, eb.adv.mem$membership)
table(eb.adv.mem$membership)
eb.adv$bridges
eb.adv$removed.edges    
eb.adv$edge.betweenness
brdg.colors <- rep('grey',length(eb.adv$removed.edges))
brdg.colors[eb.adv$removed.edges[1:5]] <- 'red'
brdg.colors[eb.adv$removed.edges[1:5]] <- 'red'
brdg.wd <- rep(1,length(eb.adv$removed.edges))
brdg.wd[eb.adv$removed.edges[1:5]] <- 5
```

## Leadership network

```{r leader}
leader <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "leadership")!=1])
deg.lead <- degree(leader,mode='in')
leader.noiso <- delete.vertices(leader,V(leader)[degree(leader)==0])
deg.lead.no <- degree(leader.noiso, mode='in')

plot(leader.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
     vertex.label=NA,vertex.color=int_vertex_colors,vertex.size=3)
plot(leader.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
     vertex.label=NA,vertex.color=prof_vertex_colors,vertex.size=3)

#plot(degree.distribution(leader))
```


# How hierarchical are the networks?

Use a measure called the *global reach centralisation*.  Based on out-ties, this is likely to be high for the full networks, as out-ties are restricted to the people interviewed (22 out of 78).

GRC for Knows: `r grc(knows)`

GRC for Advice: `r grc(advice)`

GRC for Leadership: `r grc(leader)`

Same again, but without isolates

These figures suggest a high degree of hierarchy, but this is almost certainly because only 22 of the 78 network members were interviewed, so they are the only ones with out-degrees.

GRC for Knows: `r grc(knows.noiso)`

GRC for Advice: `r grc(advice.noiso)`

GRC for Leadership: `r grc(leader.noiso)`


```{r advicerand, eval=FALSE}
# This function generates random networks with the same in and out-degree and also rewires randomly (as in the grc article).  Used to test the grc function

grc.rand <- function(g, niter=100, rwiter=100){
  indeg <- degree(g, mode='in')
  outdeg <- degree(g, mode='out')
  rand.grcs <- vector(length=niter)
  for (i in 1:niter){
    tmp <- degree.sequence.game(outdeg, indeg, method='simple')
    adv.rand <- rewire(tmp, niter=rwiter)
    rand.grcs[i] <- grc(adv.rand)
  }
  rand.grcs
}
(adv.rands <- grc.rand(advice.noiso, niter=100, rwiter=5000))

library(igraphdata)
data(macaque)
mean(degree(macaque,mode='out'))
grc(macaque)

neural <- read.graph("celegansneural.gml",format="gml")
mean(degree(neural,mode='out'))
grc(neural)
n.rnd <- grc.rand(neural)
summary(n.rnd)
```

# Use only subnetworks of people who were interviewed

First, need to generate the subnets, then plot them.

```{r subnets}
knows.int <- delete.vertices(knows,V(knows)$Interviewed=='No')
knows.deg <- degree(knows.int, mode='in')
plot(knows.int, edge.arrow.size=.5, vertex.color=prof_vertex_colors, layout=layout.fruchterman.reingold.grid, vertex.size=knows.deg, main='Knows')

advice.int <- delete.vertices(advice,V(advice)$Interviewed=='No')
adv.deg <- degree(advice.int, mode='in')
plot(advice.int, edge.arrow.size=.5, vertex.color=prof_vertex_colors, layout=layout.fruchterman.reingold.grid, vertex.size=adv.deg, main='Advice')

leader.int <- delete.vertices(leader,V(leader)$Interviewed=='No')
lead.deg <- degree(leader.int, mode='in')
plot(leader.int, edge.arrow.size=.5, vertex.color=prof_vertex_colors, layout=layout.fruchterman.reingold.grid, vertex.size=lead.deg, main='Leadership')
```

GRC for Knows: `r grc(knows.int)`

GRC for Advice: `r grc(advice.int)`

GRC for Leadership: `r grc(leader.int)`


All of these networks have no hierarchy according to the global reach centrality measure.  This should be confirmed by a dyad census, which should show many mutual ties. In each network there are 22 people, so (22^2 - 22)/2 = 231 pairs.  This is used to express numbers of types of dyad as a proportion.

```{r dyad}
know.dyad <- dyad.census(knows.int)

adv.dyad <- dyad.census(advice.int)

lead.dyad <- dyad.census(leader.int)
```

Relation | Mutual ties | Assymatric ties | Null ties
---------|-------------|-----------------|---------------------------------------------
Knows | `r know.dyad$mut` | `r know.dyad$asym` | `r know.dyad$null`
(Proportions) | `r know.dyad$mut/231` | `r know.dyad$asym/231` | `r know.dyad$null/231`
Advice | `r adv.dyad$mut` | `r adv.dyad$asym` | `r adv.dyad$null`
(Proportions) | `r adv.dyad$mut/231` | `r adv.dyad$asym/231` | `r adv.dyad$null/231`
Leadership | `r lead.dyad$mut` | `r lead.dyad$asym` | `r lead.dyad$null`
(Proportions) | `r lead.dyad$mut/231` | `r lead.dyad$asym/231` | `r lead.dyad$null/231`

As we would expect, much less reciprocity in *advice* than *knows* and less still in *leadership*. There is some evidene of status hiearchy in this, so think about other ways of investigating this. 

## Do people who give advice ask for advice from the people to whom they have given it?

```{r recipadvice, echo=FALSE}
reciprocity <- function(m){
# Katz and Power index
  ## 2 April 2013
  ## Skvoretz and Agneesens (2007) "Reciprocity, Multiplexity and Exchange Measures
  ## Quality and Quantity 41: 341-357
  require(igraph)
  if (!is.igraph(m)) stop('igraph object required')
  if (!igraph::is.directed(m)) stop('Graph must be directed')
  deg <- degree(m, mode='out')
  n <- length(deg)
  tot.tie <- sum(deg)
  mut.exp <- (tot.tie^2 - tot.tie) / (2 * (n * (n - 1) - 1))
  mut.tot <- tot.tie / 2
  mut.obs <- dyad.census(m)$mut
  (mut.obs - mut.exp) / (mut.tot - mut.exp)
}

deg.cor <- function(m){
  in.deg <- degree(m,mode='in')
  out.deg <- degree(m, mode='out')
  cor(in.deg,out.deg)
}

cyclicality <- function(m){
  # Function to caclulate cyclicality, defined as number of cycles divided by number of triads with at least 2 components of a cycle.
  # Reguires igraph graph object
  require(igraph)
  if (!is.igraph(m)) stop('igraph object required')
  if (!igraph::is.directed(m)) stop('Graph must be directed')
  tri <- triad.census(m)
  cyc <- tri[10]
  tot <- sum(tri[c(6,10)])
  cyc / tot
}

know.recip <- reciprocity(knows.int)
know.degcor <- deg.cor(knows.int)
know.cyc <- cyclicality(knows.int)

adv.recip <- reciprocity(advice.int)
adv.degcor <- deg.cor(advice.int)
adv.cyc <- cyclicality(advice.int)

lead.recip <- reciprocity(leader.int)
lead.degcor <- deg.cor(leader.int)
lead.cyc <- cyclicality(leader.int)

know.igrecip <- igraph::reciprocity(knows.int)
adv.igrecip <- igraph::reciprocity(advice.int)
lead.igrecip <- igraph::reciprocity(leader.int)
```

Correlation between in-degree and out-degree shows if people who ask lots of others for advice don't get asked for advice by a lot of people, and vice versa.

Reciprocity depends on working out the expected number of mutual ties given the total number of ties in the network.  So, this is conditioning on the number of ties (an alterative to  proportion of reciprocal ties is to total possible ties shown above).

Finally, can look at cyclicality, number of cycles as proportion of triads with at least 2 parts of the cycle.


Measure  |  Knows  |  Advice | Leadership
---------|---------|---------|------------------------------------------------------
Correlation of in- and out-degree | `r know.degcor` | `r adv.degcor` | `r lead.degcor`
Reciprocity | `r know.recip` | `r adv.recip` | `r lead.recip`
Cyclicality | `r know.cyc` | `r adv.cyc` | `r lead.cyc`
igraph Reciprocity| `r know.igrecip` | `r adv.igrecip` | `r lead.igrecip`

These measures are all quite high, and suggest that there is a high degree of mutual support in the network.  This is particularly evident in the advice network, although even people who are identified as leaders often reciprocate the identification.

# Centrality

```{r cent}
adv.bet <- betweenness(advice.int)
lead.bet <- betweenness(leader.int)
adv.cons <- constraint(advice.int)
lead.cons <- constraint(leader.int)
```
## Advice: Betweenness

`r adv.bet`

## Advice: Burt's constraint

`r adv.cons`

## Leadership: Betweenness

`r lead.bet`

## Leadership: Burt's constraint

`r lead.cons`


# Structural equivance in advice network

Another approach to investigating networks is to look for evidence of equivalent groups.  These would be groups of respondents who have similar patterns of ties to other respondents.  

### Looking just at *in* ties (get approached for advice)
```{r advicestrucin}
# Make labels for dendograms
labs <- V(advice.int)$name
lablst <- str_extract_all(labs,'[A-Z]')
labs <- sapply(lablst,function(x) paste(x[1],x[2],sep=''))


advice.jac <- similarity.jaccard(advice.int, mode='in')
advice.jac.clus <- hclust(as.dist(1-advice.jac), method='ave')
plot(advice.jac.clus, labels=labs)

clus1 <- c(5,13,14,16)
clus2 <- c(1,6,17)
clus3 <- setdiff(1:22,c(clus1,clus2))
#V(advice.int)[clus1]
#V(advice.int)[clus2]
#V(advice.int)[clus3]
deg.ad.int <- degree(advice.int, mode='in')
deg.ad.int[clus1]
deg.ad.int[clus2]
deg.ad.int[clus3]
```

This seems to show three structurally equivalent groups, the first two of which are less approached for advice than the others. (Jane Houghton and Rachel Bloomer are exceptions as they have only 9 nominations, but are in third cluster.)

### Looking just at *out* ties (approach others for advice)
```{r advicestrucout}

advice.jac.out <- similarity.jaccard(advice.int, mode='out')
advice.jac.out.clus <- hclust(as.dist(1-advice.jac.out), method='ave')
plot(advice.jac.out.clus, labels=labs)

clus1 <- c(5,13,14,16)
clus2 <- setdiff(1:22,clus1)
#V(advice.int)[clus1]
#V(advice.int)[clus2]

deg.ad.int.out <- degree(advice.int, mode='out')
deg.ad.int.out[clus1]
deg.ad.int.out[clus2]

```

People who get asked for advice least also ask for advice least.

### Looking both type of ties together
```{r advicestrucboth}

advice.jac.all <- similarity.jaccard(advice.int, mode='all')
advice.jac.all.clus <- hclust(as.dist(1-advice.jac.all), method='ave')
plot(advice.jac.all.clus, labels=labs)

clus1 <- c(5,13,14,16)
clus2 <- c(6,11)
clus3 <- setdiff(1:22,c(clus1,clus2))
 
#V(advice.int)[clus1]
#V(advice.int)[clus2]
#V(advice.int)[clus3]

deg.ad.int.all <- degree(advice.int, mode='all')
deg.ad.int.all[clus1]
deg.ad.int.all[clus2]
deg.ad.int.all[clus3]
```



## Structural equivance in leader network
### Looking just at *in* ties (nominated as leaders)

```{r leaderstrucin}
leader.jac <- similarity.jaccard(leader.int, mode='in')
leader.jac.clus <- hclust(as.dist(1-leader.jac), method='ave')
plot(leader.jac.clus, labels=labs)

clus1 <- c(1,17,19,20)
clus2 <- c(5,13)
clus3 <- c(14,16,18,22)
clus4 <- setdiff(1:22,c(clus1,clus2,clus3))
#V(leader.int)[clus1]
#V(leader.int)[clus2]
#V(leader.int)[clus3]
#V(leader.int)[clus4]

deg.lead.int <- degree(leader.int, mode='in')
deg.lead.int[clus1]
deg.lead.int[clus2]
deg.lead.int[clus3]
deg.lead.int[clus4]

```
# Regular equivalence

```{r releader, message=FALSE}
# Need to create network package object
lead.adj <- get.adjacency(leader.int, sparse=FALSE)
lead.net <- network::as.network.matrix(lead.adj)
lead.re <- sna::redist(lead.net)
lead.re.mds <- cmdscale(as.dist(lead.re))
clus1 <- c(1,2,6,8,14,22)
clus2 <- c(3,4,15,19,20,21)
clus3 <- c(5,13)
clus4 <- c(7,10,11,12,16,17)
clus5 <- 9
clus6 <- 18

plot(lead.re.mds, main='MDS of regular equivalence' )

V(leader.int)[clus1]
V(leader.int)[clus2]
V(leader.int)[clus3]
V(leader.int)[clus4]
V(leader.int)[clus5]
V(leader.int)[clus6]

colors <- RColorBrewer::brewer.pal(6,'Accent')
bgcolors <- paste(colors,'4C',sep='')
re_vertex_colors <- 1:22
re_vertex_colors[clus1 ] <- colors[1]
re_vertex_colors[clus2] <- colors[2]
re_vertex_colors[clus3] <- colors[3]
re_vertex_colors[clus4] <- colors[4]
re_vertex_colors[clus5] <- colors[5]
re_vertex_colors[clus6] <- colors[6]

plot(leader.int, mark.groups=list(clus1,clus2,clus3,clus4,clus5,clus6), edge.arrow.size=.5,vertex.color=re_vertex_colors, layout=layout.fruchterman.reingold.grid, mark.col=bgcolors,mark.border=colors, vertex.size=deg.lead.int)

 lead.eq <- sna::equiv.clust(lead.net,equiv.dist=lead.re, cluster.method='ave')
sna::plot.equiv.clust(lead.eq, labels=labs)
detach(package:network)

```

```{r leadrean, eval=FALSE}
### Plot blockmodel from regular equivalence

memb.vec <- 1:22
for (i in 1:6){
  cn <- paste('clus',1:6,sep='')
  memb.vec[get(cn[i])] <- i
}
leadbm <- sna::blockmodel(lead.net,memb.vec)
sna::plot.blockmodel(leadbm)
```

