Read network data
==============================================

```{r setup}
library(igraph)
library(stringr)
opts_chunk$set(echo=TRUE, comment='', dev='png',cache=TRUE,dpi=100,fig.width=7)
```

Find the names of the csv files in the directory.  These were saved from the original excel file.  They have the format FirstnameSecondname.csv or FirstnameSecondnameOrg.csv.
These are then read in to two lists, one with individual networks and one with organizational affiliations.

Note that the first few lines in the files do not contain valid data, and there are various notes at the bottom of the file, hence the use of `skip` and `nrows`.

```{r readcsvs}
fns <- dir(pattern='*.csv')
ix <- str_detect(fns,"Orgs")
org.fns <- fns[ix]
fns <- fns[!ix]
n <- length(fns)
nets <- list()
for (i in 1:n){
    nets[[i]] <- read.csv(fns[i],skip=5,header=FALSE,nrows=77)
}

n.org <- length(org.fns)
org.nets <- list()     
org.names <- data.frame(ID=1:52)
for (i in 1:n.org){
  org.nets[[i]] <- read.csv(org.fns[i],skip=6,header=FALSE,nrows=52)[,1:4]
  colnames(org.nets[[i]]) <- c('Organization','Patients','Staff','Other')
  org.names <- cbind(org.names,org.nets[[i]][,1])
}

```

Check all organizational names are the same in each csv file

```{r namecheck}
for (i in 3:(n.org+1)){
  print(i)
  print(all.equal(org.names[,2],org.names[,i]))
}
```

Next take the individual network list and turn it into a data frame. First column is ego name, second column is alter name (no space between first name and last name), third column is "Knows", fourth column is "Strength", fifth column is "advice", sixth column is "leadership", seventh column is "influence"

```{r dataframe}
el <- data.frame()
#i <- 1
for (i in 1:n){
  tmp <- nets[[i]][,c(1,2,3,6,7,8)]
  ego <- str_sub(fns[i],1,-5)
#   sp <- str_locate(ego,'[a-z][A-Z]')
#   ego <- paste(substr(ego,1,sp[1]) ,  substring(ego,sp[2]))
  sp2 <- str_locate(tmp[,1],' [A-Z]')
  alter <- paste(substr(tmp[,1],1,sp2[,1]-1),substring(tmp[,1],sp2[,2]),sep="")
#  ix <- tmp[,2] > 0
#  tmp <- tmp[ix,]

alter <- ifelse(alter=='RachaelWiliams','RachaelWilliams',alter)  
alter <- ifelse(alter=='SuEdwards','SueEdwards',alter)

  # Get rid of -1 s in advice variable
  tmp[,4] <- ifelse(tmp[,4]<=0, 0, 1)
  el <- rbind(el,data.frame(ego,alter,Knows=tmp[,2],Strength=tmp[,3],
                advice=tmp[,4],leadership=tmp[,5],influence=tmp[,6]))
}
```

Reduce to non-zero edges and build a graph object

```{r makegraph}
el.nonzero.edges <- subset(el, (Knows > 0 | advice > 0))


g1 <- graph.data.frame(el)
summary(g1)
```

```{r attribs}
alters <- V(g1)
el$alter %in% el$ego

attribs <- read.table('Attributes.txt',header=TRUE,stringsAsFactors=FALSE)
#attribs[,1] <- attribs[,1]-1
#names(attribs)[1] <- "Vertex"
# Set vertex attributes
for (i in V(g1)) {
  for (j in names(attribs)) {
    ix <- match(V(g1)$name[i],attribs[,1])
    if (is.na(ix)) stop("Invalid name")
    g1 <- set.vertex.attribute(g1, j, index=i, attribs[ix,j])
  }
}
summary(g1)

int_vertex_colors <- get.vertex.attribute(g1,"Interviewed")
colors <- c('Lightblue', 'Red')
int_vertex_colors[int_vertex_colors == 'No'] <- colors[1]
int_vertex_colors[int_vertex_colors == 'Yes'] <- colors[2]

knows <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "Knows")==0])


summary(knows)
deg.know <- degree(knows,mode='in')

knows.noiso <- delete.vertices(knows,V(knows)[degree(knows)==0])

x11(width=12,height=8)
plot(knows.noiso, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=5,
     vertex.label=NA,vertex.color=int_vertex_colors)
#dev.off()

advice <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "advice")==0])
summary(advice)


advice.noiso <- delete.vertices(advice,V(advice)[degree(advice)==0])
deg.adv <- degree(advice,mode='in')
plot(degree.distribution(advice.noiso))
#pdf('NetworkPlot1.pdf')
x11(width=12,height=8)
plot(advice.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
     vertex.label=NA,vertex.color=int_vertex_colors,vertex.size=3)
dev.off()

tkplot(advice.noiso, layout=layout.fruchterman.reingold.grid,edge.arrow.size=.3,
       vertex.label=NA,vertex.color=int_vertex_colors,vertex.size=deg.adv*2)
plot.coords <- tkplot.getcoords(3)
tkplot.export.postscript(3)
# Write Pajek file
write.graph(g1,'all.net',format='pajek')

deg <- degree(g1, mode='in')

plot(degree.distribution(g1))
     
bet.adv <- betweenness(advice.noiso)

x11(width=10,height=6)
plot(g1,layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.label=NA,vertex.size=deg)
dev.off()


leader <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "leadership")!=1])
deg.lead <- degree(leader,mode='in')
influence <- delete.edges(g1, E(g1)[get.edge.attribute(g1,name = "influence")!=1])


leader.noiso <- delete.vertices(leader,V(leader)[degree(leader)==0])
deg.lead.no <- degree(leader.noiso, mode='in')

x11(width=10,height=6)
tkplot(leader.noiso,layout=layout.kamada.kawai,edge.arrow.size=1,vertex.label=deg.lead.no,
     vertex.size=deg.lead.no,vertex.color=int_vertex_colors)
dev.off()

deg.inf <- degree(influence, mode='in')
infl.noiso <- delete.vertices(influence,V(influence)[degree(influence)==0])
deg.inf.no <- degree(infl.noiso, mode='in')

x11(width=10,height=6)
tkplot(infl.noiso,layout=layout.kamada.kawai,edge.arrow.size=1,vertex.label=deg.inf.no,
       vertex.size=deg.inf.no*2,vertex.color=int_vertex_colors)
dev.off()

library(xtable)
xtable(data.frame(alters$name,deg.know,deg.adv,deg.lead,deg.inf),digits=0)


xtable(data.frame(alters$name,deg.adv),digits=0)
#tk1 <- tkplot(g1,layout=layout.kamada.kawai,vertex.label=NA,vertex.size=5)

##### Focus on know network
### Lood at strength of ties
edge.colors <- c('gray','springgreen','lightskyblue','navy','red')
int_edge_colors <- get.edge.attribute(knows,"Strength")
int_edge_colors[int_edge_colors == 1] <- edge.colors[1]
int_edge_colors[int_edge_colors == 2] <- edge.colors[1]
int_edge_colors[int_edge_colors == 3] <- edge.colors[3]
int_edge_colors[int_edge_colors == 4] <- edge.colors[5]
int_edge_colors[int_edge_colors == 5] <- edge.colors[5]

# Weak ties grey
# Medium ties blue
# Strong ties red

x11(width=12,height=10)
plot(knows, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=5,
     vertex.label=deg.know,vertex.color=int_vertex_colors,edge.color=int_edge_colors)
dev.off()

# Look at advice network

tkplot(advice.noiso, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=5,
     vertex.label=deg.know,vertex.color=int_vertex_colors)

bet.adv <- betweenness(advice,directed=FALSE)
wtrp.adv <- walktrap.community(advice.noiso)
memb.adv <- community.to.membership(advice.noiso,wtrp.adv$merges,steps=12)
modularity(advice.noiso, memb.adv$membership)

vertex.colors <- c('springgreen','lightskyblue','navy','red','grey')

com.vertex.colors <- memb.adv$membership
com.vertex.colors[com.vertex.colors>3] <- vertex.colors[5]
com.vertex.colors[com.vertex.colors==0] <- vertex.colors[1]
com.vertex.colors[com.vertex.colors==1] <- vertex.colors[2]
com.vertex.colors[com.vertex.colors==2] <- vertex.colors[3]
com.vertex.colors[com.vertex.colors==3] <- vertex.colors[4]

brdg.colors <- rep('grey',199)
brdg.color[]

tkplot(advice.noiso, layout=layout.kamada.kawai,edge.arrow.size=.3,vertex.size=5,
       vertex.label=NA,vertex.color=com.vertex.colors)

ix <- order(memb.adv$membership)
com.df <- data.frame(Name=get.vertex.attribute(advice.noiso,"name"),Community=memb.adv$membership)[ix,]
com.df[1:16,]
rem <- as.numeric(rownames(com.df[17:62,])) -1
advice.com <- delete.vertices(advice.noiso,rem)
com.vertex.colors.sub <- c(1,3,2,1,0,2,2,3,2,3,2,2,2,0,0,1)
com.vertex.colors.sub[com.vertex.colors.sub==0] <- vertex.colors[1]
com.vertex.colors.sub[com.vertex.colors.sub==1] <- vertex.colors[2]
com.vertex.colors.sub[com.vertex.colors.sub==2] <- vertex.colors[3]
com.vertex.colors.sub[com.vertex.colors.sub==3] <- vertex.colors[4]

plot(advice.com, layout=layout.kamada.kawai,edge.arrow.size=1,vertex.size=5,
       vertex.label=NA,vertex.color=com.vertex.colors.sub)

## fastgreedy community structure
fg.adv <- fastgreedy.community(as.undirected(advice.noiso))
fg.adv <- community.to.membership(advice.noiso,fg.adv$merges,steps=12)
modularity(advice.noiso, fg.adv$membership)
fg.adv$membership
wtrp.adv$membership

### edge betweenness community
eb.adv <- edge.betweenness.community(advice.noiso,directed=FALSE)
eb.adv.mem <- community.to.membership(advice.noiso, eb.adv$merges, steps=9)
modularity(advice.noiso, eb.adv.mem$membership)
table(eb.adv.mem$membership)
eb.adv$bridges
eb.adv$removed.edges    
eb.adv$edge.betweenness
brdg.colors <- rep('grey',length(eb.adv$removed.edges))
brdg.colors[eb.adv$removed.edges[1:5]] <- 'red'
brdg.colors[eb.adv$removed.edges[1:5]] <- 'red'
brdg.wd <- rep(1,length(eb.adv$removed.edges))
brdg.wd[eb.adv$removed.edges[1:5]] <- 5

tkplot(advice.noiso, layout=layout.kamada.kawai,edge.arrow.size=1,vertex.size=5,
     edge.color=brdg.colors,
     edge.width=brdg.wd)

##### Network of people interviewed only

V.notint <- V(knows)[V(knows)$Interviewed=='Yes']
net.interview <- delete.vertices(knows,V(knows)[V(knows)$Interviewed=='No'])
summary(net.interview)
dens <- function(g){
  n.edge <- ecount(g)
  n.vert <- vcount(g)
  e.max <- n.vert^2 - n.vert
  if (is.directed(g)){
    n.edge/e.max
  }
  else {
    n.edge/(e.max/2)
  }
}

dens(net.interview)
```